Red-Giant Hot Jupiters: Brilliant Radio Beacons

### 2014-03-29 ms.tex
added my name
recent change

\author{
%
David S. Spiegel\altaffilmark{1} \\
%
and some order: \\
%
Nikku Madhusudhan\altaffilmark{?} [Add correct affiliation] \\
%
Mehrdad Mirbabayi\altaffilmark{1} \\
%
Aaron Parsons\altaffilmark{?} [Add correct affiliation] \\
%
Tony Mroczkowski\altaffilmark{?} [Add correct affiliation]
}

@@

\begin{abstract}
  Red-giant hot Jupiters are jovian planets orbiting red-giant-branch
  or asymptotic-giant-branch (AGB) stars.  Post-main-sequence stars
  lose mass at much higher rates than main-sequence stars.  A jovian
  planet passing through the dense winds of its AGB host can capture
  stellar wind in its magnetosphere.  The cyclotron frequency of
  electrons from the stellar wind accreting onto the planet scales as
  100~MHz~$(B/3 {\rm~Gauss})$.  Such a planet might generate a radio
  luminosity that would be visible from kiloparsec distances.
\end{abstract}

@@

% fix
Roughly 20\% of the more than 700 currently known
exoplanets\footnote{See online catalogs such as
  http://www.openexoplanetcatalogue.com/ \citep{rein2012},
  http://exoplanet.eu \citep{schneider_et_al2011}, or
  http://exoplanets.org \citep{wright_et_al2011} for up-to-date
  lists.} have masses greater than half of Jupiter's, orbital radii
greater than 1~AU, and will become hot Jupiters (i.e., for the present
purposes, this means they will receive at least as much irradiation as
the hot Neptune

The cyclotron frequency is
\begin{equation}
f_{\rm cyclotron} = \frac{eB}{2\pi m_e c} \approx 3 {\rm~MHz} \left( \frac{B}{1 \rm~G} \right) \, .
\label{eq:cyc}
\end{equation}
For magnetic field strengths in the vicinity of $\sim$25-50~Gauss (G),
the cyclotron frequency will be in the range $\sim$75-150~MHz.

@@

Noting that
\begin{equation}
\dot{M}_* = 4\pi r^2 \rho[r] v \, ,
\label{eq:mdot
}\end{equation}
so $\rho[r] = \dot{M}_*/(4\pi r^2 v)$,
\begin{eqnarray}
\frac{\rho v^2}{2} = \frac{B^2}{8\pi} \sim \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-3} \, ,
\end{eqnarray}
where $B_0$ is the field strength at a distance $d_0$ from the planet.
\begin{eqnarray}
\frac{\rho v^2}{2} & \sim & \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-6} \\
\frac{\dot{M}_* v}{8\pi a^2} & = & \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-6} \\
\frac{\dot{M}_* v}{r^2} & = & B_0^2 \left( \frac{d}{d_0} \right)^{-6} \\
d^6 & = & \frac{d_0^6 B_0^2 r^2}{\dot{M}_* v} \\
d_A & = & d_0 \left( \frac{B_0^2 r^2}{\dot{M}_* v} \right)^{1/6} \\
  & \sim & 4 R_J \left( \frac{d_0}{R_J}\right) \left\{ \frac{\left( \frac{B}{10 \rm~G} \right)^2 \left( \frac{r}{5 \rm~AU} \right)^2}{\left( \frac{\dot{M}_*}{10^{-6} M_\odot/\rm yr} \right) \left( \frac{v}{20 \rm~km/s} \right)} \right\}^{1/6}
\label{eq:Chapman-Ferraro}
\end{eqnarray}
In the above, $d_A$ is the distance away from the planet where the
magnetic energy density equals the kinetic energy density in the
stellar wind, if $B_0$ is the magnetic field strength at distance
$d_0$.

@@

\bibliography{biblio.bib}

>>>>

\author{
%
David S. Spiegel\altaffilmark{1} \\
%
{\bf and some order:} \\
%
Nikku Madhusudhan\altaffilmark{?} [Add correct affiliation] \\
%
Mehrdad Mirbabayi\altaffilmark{1} \\
%
Aaron Parsons\altaffilmark{?} [Add correct affiliation] \\
%
Tony Mroczkowski\altaffilmark{?} [Add correct affiliation]
}

@@

\begin{abstract}
  Red-giant hot Jupiters are jovian planets orbiting red-giant-branch
  or asymptotic-giant-branch (AGB) stars.  Post-main-sequence stars
  lose mass at much higher rates than main-sequence stars.  A jovian
  planet passing through the dense winds of its AGB host can capture
  stellar wind in its magnetosphere.  The cyclotron frequency of
  electrons from the stellar wind accreting onto the planet scales as
  100~MHz~$(B/30 {\rm~Gauss})$.  Such a planet might generate a radio
  luminosity that would be visible from kiloparsec distances.
\end{abstract}

@@

% fix
When stars less than $\sim$8~$M_\sun$ evolve off the main sequence,
they go through stages on the red-giant branch (RGB) and the
asymptotic-giant branch (AGB) when their radii and luminosities
increase by orders of magnitude.  Jovian planets in orbit around such
stars can be transiently heated to hot-Jupiter temperatures
($\sim$1000~K or more) at distances out to tens of AU, depending on
the star's mass.  These planets' existence was predicted by
\citet{spiegel+madhusudhan2012} and we will refer to them as
``red-giant hot Jupiters'' (RGHJs).

Roughly 20\% of the more than 700 currently known
exoplanets\footnote{See online catalogs such as
  http://www.openexoplanetcatalogue.com/ \citep{rein2012},
  http://exoplanet.eu \citep{schneider_et_al2011}, or
  http://exoplanets.org \citep{wright_et_al2011} for up-to-date
  lists.} have masses greater than half of Jupiter's, orbital radii
greater than 1~AU, and will become hot Jupiters (i.e., for the present
purposes, this means they will receive at least as much irradiation as
the hot Neptune

The cyclotron frequency is
\begin{equation}
f_{\rm cyclotron} = \frac{eB}{2\pi m_e c} \approx 2.8 {\rm~MHz} \left( \frac{B}{1 \rm~G} \right) \, .
\label{eq:cyc}
\end{equation}
For magnetic field strengths in the vicinity of $\sim$25-50~Gauss (G),
the cyclotron frequency will be in the range $\sim$75-150~MHz.

@@

Noting that
\begin{equation}
\dot{M}_* = 4\pi r^2 \rho[r] v \, ,
\label{eq:mdot
}\end{equation}
so $\rho[r] = \dot{M}_*/(4\pi r^2 v)$,
\begin{eqnarray}
\frac{\rho v^2}{2} = \frac{B^2}{8\pi} \sim \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-3} \, ,
\end{eqnarray}
where $B_0$ is the field strength at a distance $d_0$ from the planet.
\begin{eqnarray}
\frac{\rho v^2}{2} & \sim & \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-6} \\
\frac{\dot{M}_* v}{8\pi a^2} & = & \frac{B_0^2}{8\pi} \left( \frac{d}{d_0} \right)^{-6} \\
\frac{\dot{M}_* v}{r^2} & = & B_0^2 \left( \frac{d}{d_0} \right)^{-6} \\
d^6 & = & \frac{d_0^6 B_0^2 r^2}{\dot{M}_* v} \\
d_A & = & d_0 \left( \frac{B_0^2 r^2}{\dot{M}_* v} \right)^{1/6} \\
  & \sim & 4 R_J \left( \frac{d_0}{R_J}\right) \left\{ \frac{\left( \frac{B}{10 \rm~G} \right)^2 \left( \frac{r}{5 \rm~AU} \right)^2}{\left( \frac{\dot{M}_*}{10^{-6} M_\odot/\rm yr} \right) \left( \frac{v}{20 \rm~km/s} \right)} \right\}^{1/6}
\label{eq:Chapman-Ferraro}
\end{eqnarray}
In the above, $d_A$ is the distance from the planet to the Alfven
point, where the magnetic energy density $u_B = B^2 / 8\pi$ equals the
kinetic energy density in the stellar wind $u_w = \rho v^2/2$, and
$B_0$ is the magnetic field strength at distance $d_0$.

@@

mehrdad
\newpage
\bibliography{biblio.bib}
